Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-07-26 TEX/draft4.tex
כובל ופ רקבםהקרט

\TODO{Frank: Working with Owen Saxton, I was able to show that alignment via cross-correlation would be accurate enough for the purpose of aligning noisy images of single molecules if the dose exceeded a certain level, which depended on the molecule?s size and contrast against the background.[19a] Using the ribosome as an example, it became clear from the formula we obtained that the single-particle approach to structure research was indeed feasible for molecules of sufficient size: Particle Size > 3/[Contrast$^2$ x Resolution (as length) X Critical Electron Dose] A detailed examination of this dependency was later to be undertaken by Richard Henderson.[19b]. From Joachim Frank?s Nobel Lecture. Amit's email from July 10.}
In that paper and in others that followed (e.g.,~\cite{glaeser1999electron}), it was established that particle picking is impossible for molecules below a certain weight.
%This has often been understood as meaning that particle reconstruction itself is impossible for such molecules~\cite{khoshouei2017cryo,liu2018nearatomic,liang2017phase}. \TODO{This list may be unfair, as some of these papers (in particular, \cite{liang2017phase}) just state that small molecules are very challenging, which is true.}
As a result, it is impossible to reconstruct such small molecules by any of the existing computational pipelines for single particle analysis in cryo--EM, as the particles themselves cannot be picked from the micrographs.
%
%\TODO{either here, or in a close neighborhood of this paragraph you should mention the advances that the field has made in recent years towards reconstruction of small size molecules: mention Volta phase plate and the scaffolding cage (and cite the papers I sent you).}
This has motivated recent technical advances in the field, including the use of Volta phase plates~\cite{khoshouei2017cryo,liang2017phase} and scaffolding cages~\cite{liu2018nearatomic}.
%
%
Despite this progress, detecting small molecules in the micrographs (below $\sim$50 kDa) remains a challenge.
We note that nuclear magnetic resonance (NMR) spectroscopy and X-ray crystallography are well suited to reconstruct small molecules. Yet, cryo--EM has a lot to offer even for molecules with already known structures obtained via NMR spectroscopy or X-ray crystallography, because these methods have limited ability to distinguish conformational variability. \TODO{Need a ref for this claim.}

>>>>

\TODO{Frank: Working with Owen Saxton, I was able to show that alignment via cross-correlation would be accurate enough for the purpose of aligning noisy images of single molecules if the dose exceeded a certain level, which depended on the molecule?s size and contrast against the background.[19a] Using the ribosome as an example, it became clear from the formula we obtained that the single-particle approach to structure research was indeed feasible for molecules of sufficient size: Particle Size > 3/[Contrast$^2$ x Resolution (as length) X Critical Electron Dose] A detailed examination of this dependency was later to be undertaken by Richard Henderson.[19b]. From Joachim Frank's Nobel Lecture. Amit's email from July 10. Btw, it is also on slide 16 of his lecture}

In that paper and in others that followed (e.g.,~\cite{glaeser1999electron}), it was established that particle picking is impossible for molecules below a certain weight.
%This has often been understood as meaning that particle reconstruction itself is impossible for such molecules~\cite{khoshouei2017cryo,liu2018nearatomic,liang2017phase}. \TODO{This list may be unfair, as some of these papers (in particular, \cite{liang2017phase}) just state that small molecules are very challenging, which is true.}
As a result, it is impossible to reconstruct such small molecules by any of the existing computational pipelines for single particle analysis in cryo--EM, as the particles themselves cannot be picked from the micrographs.
%
%\TODO{either here, or in a close neighborhood of this paragraph you should mention the advances that the field has made in recent years towards reconstruction of small size molecules: mention Volta phase plate and the scaffolding cage (and cite the papers I sent you).}
This has motivated recent technical advances in the field, including the use of Volta phase plates~\cite{khoshouei2017cryo,liang2017phase} and scaffolding cages~\cite{liu2018nearatomic}.
%
%
Despite this progress, detecting small molecules in the micrographs (below $\sim$50 kDa) remains a challenge.
We note that nuclear magnetic resonance (NMR) spectroscopy and X-ray crystallography are well suited to reconstruct small molecules. Yet, cryo--EM has a lot to offer even for molecules with already known structures obtained via NMR spectroscopy or X-ray crystallography, because these methods have limited ability to distinguish conformational variability. \TODO{Need a ref for this claim.}

### 2018-07-26 TEX/draft_biology.tex
כובל ופ רקבםהקרט

\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\SUM}{\text{sum}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\II}{\mathcal{I}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\mb}{\mathbf}

@@

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\hx}{\hat{x}} 
\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{\ensuremath{\textsf{SNR}}}

@@

Let $x$ be the sought image and let $y\in\RNN$ be the observed micrograph (notice that it is equivalent to think of the data as being one long micrograph or multiple smaller micrographs concatenated into one.)
The forward model (or ``image'' formation model) is as follows. An unknown binary signal  indicates (with 1's) the starting positions of all occurrences of $x$ in $y$, so that, with additive white Gaussian noise:
\begin{align}
	y & =  x \ast s + \varepsilon, & \varepsilon & \sim \mathcal{N}(0,\sigma^2 I_N),
	\label{eq:model}
\end{align}
where $\ast$ denotes linear convolution. Let $\textbf{i}:=(i_1,i_2)$ and similarly for \textbf{j}.
The binary signals obey the following property:
\begin{align}
	\textrm{If } s[\textbf{i}] = 1 \textrm{ then } s[\textbf{j}] = 0 \textrm{ for all}  \|\textbf{i}-\textbf{j}\|_\infty\leq 2L-1.
	\label{eq:spacing}
\end{align}
In words: the starting positions of any two occurrences must be separated by at least $2L-1$ positions in each direction, so that their end points are necessarily separated by at least $L-1$ signal-free entries in the micrograph.

@@

%\TODO{$L$ is a bad notation here since it could be L or n; and k is poor notation too because we use it to sum up to K.}
%
%In general, for a signal $z$ of length $m$, the autocorrelation of order $q \geq 1$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
%\begin{align}
%	a_z^q[\ell_1,\ldots,\ell_{k-1}]  & = \frac{1}{m} \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
%	\label{eq:ac_general}
%\end{align}
%where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
%For our purposes, this will be applied both to $x_k$'s (each of length $L$) and to $y$ (of length $N$).
For the purpose of this experiment, we need only the mean of the signal and its second-order autocorrelation defined for an arbitrary signal $z\in\mathbb{R}^{m\times m}$ by \TODO{to verify}
 \begin{align} 
 a_z^2[\ell_1,\ell_2] & = \frac{1}{m^2} \sum_{i_1 = \max\{0, -\ell_1\}}^{m-1 + \min\{0, -\ell_1\}} \sum_{i_2 = \max\{0, -\ell_2\}}^{m-1 + \min\{0, -\ell_2\}}z[i_1,i_2]z[i_1+\ell_1,i_2+\ell_2]. 
 \end{align}
 It can be shown that in the limit of $N\to\infty$,  the mean of the microgaph is equal to the mean of the signal $x$, scaled by the scalar 
 \begin{align}
 \gamma & = \frac{M L}{N},
 \end{align}
 which denotes the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The spacing constraint~\eqref{eq:spacing} imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.
 Similarly, the second-order autocorrelations are related through 
 \begin{align*}
	\lim_{N\to\infty} a_y^2[\ell_1,\ell_2] & = \gamma a_{x}^2[\ell_1,\ell_2] + \sigma^2\delta[\ell_1,\ell_2].
 \end{align*}
 These relations, and their extension to multiple planted images, are proven and discussed in~\cite{bendory2018estimation}.

@@

\subsection{3-D experiment}

We derive a relation between the steered bispectrum derived in Sect.~\ref{sec:steering} to the volume, expanded in a suitable basis. Specifically, expand the Fourier-transformed volume as
\[ \widehat V(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell,m}(\kk/k),\]
for $k\leq 1$ and zero otherwise, where $c$ is the assumed bandlimit, $Y_{\ell,m}$ may be taken to be either real or complex, and $j_{\ell,s}(r)$ is some radial basis function (in practice, either spherical bessel or the radial part of the 3D PSWFs as in [Lederman]). We work here with the complex spherical harmonics, given by
\[ Y_{\ell,m}(\theta,\varphi) = \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
where $P_{\ell}^m$ are the associated Legendre polynomials with the Condon-Shortley phase.

\TODO{To write down the definition of the $j$ --- the spherical bessel (to copy from the ISBI)}

In this experiment, the micrograph is related to the projections by 
\begin{equation}
\mathcal{I}(\rr) = \sum_{i=1}^{N} I_{\omega_i}(\rr-\rr_i)+ \varepsilon, \quad \varepsilon~\sim\mathcal{N}(0,\sigma^2 I),
\end{equation}
where $\rr$ is a vector in $\mathbb{R}^2$. \TODO{To write down the definition of projection}

In this paper, we need the first three autocorrelations. In particular, it can be shown that 
\begin{align}
m_k[\II]  \to \gamma\langle m_k[I_\omega]\rangle_{\omega}, 
\end{align}
where $\langle\cdot\rangle$ denotes averaging. \TODO{Do we want to define a general moment?}

In a standard SPR experiment, we expect the distribution of in-plane rotations to be uniform because .... 
Therefore, based on~\cite{landa2017steerable}, we expand the each autocorrelation functions in Prolates .... as discussed in Appendix TKTK.

The next stage is to connect the Prolates coefficients of the moments with those of the volume.
In Appendix TKTK we show that the volume is related to autocorrelations of the micrograph through

\begin{align}
\langle m_1[I_\omega]\rangle_\omega  &= \frac{L}{\sqrt{4\pi}}\sum_sa_{0,0,s}j_{0,s}(0) \\ 
\mathfrak{m}_2(q) &= \frac{\sqrt{2\pi}}{4\pi}\sum_{\substack{\ell,m\\s_1,s_2}}a_{\ell,m,s_1}\overline{a_{\ell,m,s_2}}
\int_0^1j_{\ell,s_1}(k)j_{\ell,s_2}(k)R_{0,q}(k)k\, dk \\
\mathfrak{m}_3(k,q_1,q_2) &= \sum_{\substack{\ell_1,m_1,s_1\\\ell_2,m_2,s_2\\s_3}}\sum_{\ell_3=|\ell_1-\ell_2|}^{\min(L,\ell_1+\ell_2)}C a_{\ell_1,m_1,s_1}a_{\ell_2,m_2,s_2}\overline{a_{\ell_3,m_1+m_2,s_3}}
\end{align}
where
\begin{align*}
C&:= (-1)^{m_1+m_2}\left(\begin{array}{ccc}\ell_1 & \ell_2  & \ell_3\\ m_1 & m_2 & -m_1-m_2\end{array}\right)\\
&\times \sum_{N_1=-\ell_1}^{\ell_1}\sum_{N_2=-\ell_2}^{\ell_2}(-1)^{N_1+N_2}\left(\begin{array}{ccc}\ell_1 & \ell_2  & \ell_3\\ N_1 & N_2 & -N_1-N2\end{array}\right)\int_{\rr}\Psi_{\ell_1,N_1,s_1}(\rr)\rho_{\ell_2,N_2,s_2}^{(k,q_1)}(\rr)\overline{\rho_{\ell_3,N_1+N_2,s_3}^{(k,q_2)}(\rr)},
\end{align*}
and 
\begin{align*}
\rho_{\ell,N,s}^{(k,q)}&=\int_{\Delta\rr}\Psi_{\ell,N,s}(\rr+\Delta\rr)\overline{\psi_{k,q}(\Delta\rr)}.
\end{align*}

From the data we can estimate (see Appendix TKTK) 
\begin{align*}
\widetilde{\mathfrak{m}}_2(q) &= \mathfrak{m}_2(q) + \sigma^2\frac{1}{\sqrt{2\pi}}R_{0,q}(0) \\
\widetilde{\mathfrak{m}}_3(k,q_1,q_2) &= \mathfrak{m}_3(k,q_1,q_2) +
\sigma^2m_1[\II]\left[\delta_{q_1,q_2} + \delta_{k,0}\frac{1}{2\pi}(\alpha_{0,q_1}+\alpha_{0,q_2})R_{0,q_1}(0)R_{0,q_2}(0)\right].
\end{align*}

@@

\TODO{To rewrite for the discrete case}
We consider the moments of $\II$ in the limit $N\to\infty$. To take this limit, we shall assume that $N=\Omega(M^2)$ (so by the above argument, in fact $N=\Theta(M^2)$), so $\gamma = \lim_{N\to\infty}\frac{N}{M^2}\in(0,1)$. We also assume that the orientations of the projections in $\II$ are uniformly distributed over the sphere. For the first moment,
\[ m_1[\II] =
\frac{NL^2}{M^2}\cdot\frac{1}{N}\sum_{i=1}^N\frac{1}{L^2}\int_{\RR^2}I_{\omega_i}(\rr)\,
d\rr \to \gamma\int_{\text{SO}(3)} L^2m_1[I_{\omega}]\, d\omega =
\gamma L m_1[V],\]
since $\int_{\RR^2} I_{\omega}(x,y)\, dx\, dy = \widehat I_{\omega}(\mathbf{0})=\widehat V(\mathbf{0})$ as the Fourier transform of any projection corresponds to a 2D slice of the Fourier transform of the volume passing through the origin.

For the second and third moments, note that if $\rr\in S_i$ for some $i$, then $\rr+\Delta\rr\in S_i\cup Z$ whenever $||\Delta\rr||\leq 2L$ by the separation assumption $d> 2L$. Since $\RR^2=Z\sqcup\bigsqcup_{i=1}^NS_i$, and $\II(\rr)=0$ if $\rr\in Z$, we have for $||\Delta\rr||\leq 2L$
\[\begin{aligned} m_2[\II](\Delta \rr)  &= \frac{1}{M^2}\int_{\RR^2}\II(\rr)\II(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{1}{M^2}\int_Z\II(\rr)\II(\rr+\Delta\rr)\, d\rr + \frac{1}{M^2}\sum_{i=1}^N\int_{S_i}\II(\rr)\II(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{1}{M^2}\sum_{i=1}^N\int_{\RR^2}I_{\omega_i}(\rr)I_{\omega_i}(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{N}{M^2}\cdot\frac{1}{N}\sum_{i=1}^NL^2m_2[I_{\omega_i}](\Delta\rr)\\
&\to \gamma\int_{\text{SO}(3)}L^2m_2[I_{\omega}](\Delta\rr)\, d\omega.\end{aligned}\]
If we set $m_2(\Delta \rr) = 0$ for $||\Delta\rr||>2L$, the above equality holds for all $\Delta\rr$, and taking its Fourier transform (with respect to $\Delta\rr$) and interchanging the Fourier integral with the integral over SO(3) (which can \emph{always} be done), we get
\[ \widehat{m_2[\II]}(\kk) \to \gamma\int_{\text{SO}(3)}L^2\widehat{m_2[I_{\omega}]}(\kk)\, d\omega = \gamma\langle|\widehat I_{\omega}(\kk)|^2\rangle_{\omega},\]
where $\langle\cdot\rangle_{\omega}$ denotes average over the orientations, so over SO(3).

@@

We derive a relation between the steered bispectrum derived in Sect.~\ref{sec:steering} to the volume, expanded in a suitable basis. Specifically, expand the Fourier-transformed volume as
\[ \widehat V(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell,m}(\kk/k),\]
for $k\leq 1$ and zero otherwise, where $c$ is the assumed bandlimit, $Y_{\ell,m}$ may be taken to be either real or complex, and $j_{\ell,s}(r)$ is some radial basis function (in practice, either spherical bessel or the radial part of the 3D PSWFs as in [Lederman]). We work here with the complex spherical harmonics, given by
\[ Y_{\ell,m}(\theta,\varphi) = \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
where $P_{\ell}^m$ are the associated Legendre polynomials with the Condon-Shortley phase.

@@

>>>>

\newcommand{\LL}{\mathcal{L}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\ep}{\varepsilon}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\GCD}{\mathbf{GCD}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\SUM}{\text{sum}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\II}{\mathcal{I}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\mb}{\mathbf}
\newcommand{\mk}{\mathfrak}

@@

\newcommand{\reals}{\mathbb{R}}
\newcommand{\RL}{\mathbb{R}^L}
\newcommand{\CL}{\mathbb{C}^L}
\newcommand{\RN}{\mathbb{R}^N}
\newcommand{\RNN}{\mathbb{R}^{N\times N}}
\newcommand{\CNN}{\mathbb{C}^{N\times N}}
\newcommand{\inner}[1]{\left\langle {#1} \right\rangle}
\newcommand{\hx}{\hat{x}} 
\newcommand{\one}{\mathbf{1}} 
\newcommand{\SNR}{\ensuremath{\textsf{SNR}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}

@@

Let $x$ be the sought image and let $y\in\RNN$ be the observed micrograph (notice that it is equivalent to think of the data as being one long micrograph or multiple smaller micrographs concatenated into one.)
The forward model (or ``image'' formation model) is as follows. An unknown binary signal  indicates (with 1's) the starting positions of all occurrences of $x$ in $y$, so that, with additive white Gaussian noise:
\begin{align}
	y & =  x \ast s + \varepsilon, & \varepsilon & \sim \mathcal{N}(0,\sigma^2 I_N),
	\label{eq:model}
\end{align}
where $\ast$ denotes linear convolution. Let $\textbf{i}:=(i_1,i_2)$ and similarly for \textbf{j}.
The binary signals obey the following property: \TODO{Rewrite in terms of $\delta$ functions for consistency with 3D}
\begin{align}
	\textrm{If } s[\textbf{i}] = 1 \textrm{ then } s[\textbf{j}] = 0 \textrm{ for all}  \|\textbf{i}-\textbf{j}\|_\infty\leq 2L-1.
	\label{eq:spacing}
\end{align}
In words: the starting positions of any two occurrences must be separated by at least $2L-1$ positions in each direction, so that their end points are necessarily separated by at least $L-1$ signal-free entries in the micrograph.

@@

%\TODO{$L$ is a bad notation here since it could be L or n; and k is poor notation too because we use it to sum up to K.}
%
%In general, for a signal $z$ of length $m$, the autocorrelation of order $q \geq 1$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
%\begin{align}
%	a_z^q[\ell_1,\ldots,\ell_{k-1}]  & = \frac{1}{m} \sum_{i=-\infty}^{+\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}],
%	\label{eq:ac_general}
%\end{align}
%where indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
%For our purposes, this will be applied both to $x_k$'s (each of length $L$) and to $y$ (of length $N$).
For the purpose of this experiment, we need only the mean of the signal and its second-order autocorrelation defined for an arbitrary signal $z\in\mathbb{R}^{m\times m}$ by \TODO{to verify}
 \begin{align} 
 a_z^2[\ell_1,\ell_2] & = \frac{1}{m^2} \sum_{i_1 = \max\{0, -\ell_1\}}^{m-1 + \min\{0, -\ell_1\}} \sum_{i_2 = \max\{0, -\ell_2\}}^{m-1 + \min\{0, -\ell_2\}}z[i_1,i_2]z[i_1+\ell_1,i_2+\ell_2]. 
 \end{align}
 It can be shown that in the limit of $N\to\infty$,  the mean of the microgaph is equal to the mean of the signal $x$, scaled by the scalar 
 \begin{align}
 \gamma & = \frac{M L^2}{N^2},
 \end{align}
 which denotes the density of $x$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $x$.) The spacing constraint~\eqref{eq:spacing} imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$.
 Similarly, the second-order autocorrelations are related through 
 \begin{align*}
	\lim_{N\to\infty} a_y^2[\ell_1,\ell_2] & = \gamma a_{x}^2[\ell_1,\ell_2] + \sigma^2\delta[\ell_1,\ell_2].
 \end{align*}
 These relations, and their extension to multiple planted images, are proven and discussed in~\cite{bendory2018estimation}.

@@

\subsection{3-D experiment}
\TODO{Add similarities to 2-D}

\TODO{Proposal: Use W for size of projection, M for size of micrograph}

Let $\phi$ be the Coulomb potential representing the molecule we wish to recover, and denote it 3-D Fourier transform by $\widehat \phi$. In this paper, we assume that $\widehat\phi$ is given as a finite expansion of the form
\be\label{eq:volume_expansion} 
\widehat \phi(k, \theta, \varphi) = \sum_{\ell = 0}^L\sum_{m=-\ell}^{\ell}\sum_{s=1}^{S(\ell)}a_{\ell, m, s}Y_{\ell}^m(\theta,\varphi)j_{\ell,s}(k),
\ee
where $Y_{\ell}^m$ are complex spherical harmonics and $j_{\ell,s}$ are normalized spherical Bessel functions, see Appendix TKTK for definitions. \TODO{Mention somewhere that since $\phi$ is real-valued, we recover only coefficients with $m\geq 0$.} Given $\omega\in\text{SO}(3)$, denote by $I_{\omega}$ the projection obtained from viewing direction $\omega$, and denote by $\widehat I_{\omega}$ its 2-D Fourier transform. By the Fourier projection-slice theorem~\cite[p. 11]{natterer1986mathematics}, the 2D Fourier transform of $I_{\omega}$ is given by \TODO{Move to appendix?}
\be\label{eq:projection_model}
\widehat I_{\omega}(k,\varphi) = \sum_{\ell,m,m',s}a_{\ell,m,s}D_{m',m}^{\ell}(\omega)Y_{\ell}^{m'}\left(\frac{\pi}{2},\varphi\right)j_{\ell,s}(k),\ee
where $D_{m',m}^{\ell}(\omega)$ is a Wigner-D matrix rotating spherical harmonics.

Let $\II\in\RR^{M\times M}$ denote a micrograph. For this paper, we assume that the micrograph consists of shifted copies of projections and perturbed by additive white Gaussian noise:
\begin{equation}\label{eq:micrograph_model}
\II = \sum_{j=1}^{N} I_{\omega_j}\ast\delta_{\mb s_j}+ \varepsilon, \quad \varepsilon~\sim\mathcal{N}(0,\sigma^2 I),
\end{equation}
where the viewing directions $\omega_j$ are uniformly distributed over $\text{SO}(3)$, $\mb s_j$ denotes the location of the center of the $j$th projection in the micrograph, and we impose a separation condition similarly to Eq.~\TODO{2D}:
\be\label{eq:sep_cond} ||\mb s_j - \mb s_i||_{\infty} \geq 2L. \ee
\TODO{Say something about nonuniformity in practice? Remember that biologists will jump at this...}
Define the $k$th autocorrelation of $\II$ as
\be\label{eq:Kth_autocorrelation} 
m_k[\II](\Delta \mb i_1,\ldots, \Delta \mb i_{k-1}) = \frac{1}{M^2}\sum_{i,j = 1}^M\II(\mb i)\II(\mb i+\Delta\mb i_1)\cdots \II(\mb i+\Delta \mb i_{k-1}),
\ee
where $\mb i = (i,j)$, $\Delta \mb i = (\Delta i, \Delta j)$, we require $||\Delta\mb i||_{\infty}\leq 2L-1$, and we set $\II(i,j)=0$ if either $i,j\notin\{1,\ldots, M\}$. In our procedure, we compute the first three autocorrelations, called the mean, power spectrum, and bispectrum, of the micrographs. \TODO{This definition and terminology should come earlier}. Because of the separation condition Eq.~(\ref{eq:sep_cond}), the autocorrelations of the micrograph are related to those of the projections by 
\begin{align}
m_k[\II]  \to \gamma\langle m_k[I_\omega]\rangle_{\omega} + b_k, 
\end{align}
where $\langle\cdot\rangle$ denotes averaging, $b_k$ is a bias term, and we take the limit as the number of projections $N$, and hence the dimensions of the micrograph $M$, tend to $\infty$. The mean is unbiased so $b_1 = 0$, the power spectrum has bias $b_2$ depending only on $\sigma^2$, the variance of the noise, and the bias of the bispectrum $b_3$ depends additionally on the mean of $\phi$, which can be accurately estimated from the mean of the micrograph. We can therefore estimate the quantities $\gamma\langle m_k[I_{\omega}]\rangle_{\omega}$ directly from the micrograph.

Since we assume that the viewing directions $\omega_j$ are uniformly distributed over $\text{SO}(3)$, we would like to average over all in-plane rotations of the projections in the computation of the autocorrelations. To do so, we follow~\cite{landa2017steerable,zhao2016fast} and expand the second and third autocorrelations of the micrograph in Prolate Spheroidal Wave Functions (PSWFs). 
It can then be shown that the expansion coefficients $\mathfrak{m}_k$ of the autocorrelations in PSWFs are related to the expansion coefficients of the volume by

\begin{align}\label{eq:coeffs_to_moms}
&\mathfrak{m}_1\{a_{\ell,m,s}\} = \frac{L}{\sqrt{4\pi}}\sum_sa_{0,0,s}j_{0,s}(0) \\ 
&\mathfrak{m}_2\{a_{\ell,m,s}\}(q) = \frac{\sqrt{2\pi}}{4\pi}\sum_{\substack{\ell,m\\s_1,s_2}}C_2^{(q)}(\ell,s_1,s_2)a_{\ell,m,s_1}\overline{a_{\ell,m,s_2}}
 \\
&\mathfrak{m}_3\{a_{\ell,m,s}\}(k,q_1,q_2) = \sum_{\substack{\ell_1,m_1,s_1\\\ell_2,m_2,s_2\\s_3}}\sum_{\ell_3=|\ell_1-\ell_2|}^{\min(L,\ell_1+\ell_2)}C_3^{(k,q_1,q_2)}(\ell_1,m_1,s_1;\ell_2,m_2,s_2;\ell_3,s_3)a_{\ell_1,m_1,s_1}a_{\ell_2,m_2,s_2}\overline{a_{\ell_3,m_1+m_2,s_3}},
\end{align}
see Appendix TKTK for the definitions of $C_2,C_3$.

To recover the expansion coefficients $\{a_{\ell,m,s}\}$ from the autocorrelation coefficients estimated from the micrograph $\mk m_k[\II]$, we minimize the weighted least-squares problem
\be\label{eq:min_problem_cryo} \min_{\{a_{\ell,m,s}\}} \sum_{k=1}^3\frac{\Big|\Big|\mk m_k\{a_{\ell,m,s}\} - \mk m_k[\II]\Big|\Big|_F^2}{||\mk m_k[\II]||_F^2}.\ee

@@

\TODO{To rewrite for the discrete case}

We consider the moments of $\II$ in the limit $N\to\infty$. To take this limit, we shall assume that $N=\Omega(M^2)$, and that $\gamma = \lim_{N\to\infty}\frac{N}{M^2}\in(0,1)$ is a constant. 
Because of the separation condition Eq.~(\ref{eq:sep_cond}), if $\mb i$ is in the support of the $j$th projection, then $\mb i + \Delta\mb i$ is either in the support of the same projection or outside the support of any projection. Formally, $\mb i$ is in the support of the $j$th projection if and only if $\mb 0\leq \mb i - \mb s_j < (L,L)^T$ where the inequalities apply to each coordinate. Then because of the separation condition,

note that if $\rr\in S_i$ for some $i$, then $\rr+\Delta\rr\in S_i\cup Z$ whenever $||\Delta\rr||\leq 2L$ by the separation assumption $d> 2L$. Since $\RR^2=Z\sqcup\bigsqcup_{i=1}^NS_i$, and $\II(\rr)=0$ if $\rr\in Z$, we have for $||\Delta\rr||\leq 2L$
\[\begin{aligned} m_2[\II](\Delta \rr)  &= \frac{1}{M^2}\int_{\RR^2}\II(\rr)\II(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{1}{M^2}\int_Z\II(\rr)\II(\rr+\Delta\rr)\, d\rr + \frac{1}{M^2}\sum_{i=1}^N\int_{S_i}\II(\rr)\II(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{1}{M^2}\sum_{i=1}^N\int_{\RR^2}I_{\omega_i}(\rr)I_{\omega_i}(\rr+\Delta\rr)\, d\rr\\ 
&= \frac{N}{M^2}\cdot\frac{1}{N}\sum_{i=1}^NL^2m_2[I_{\omega_i}](\Delta\rr)\\
&\to \gamma\int_{\text{SO}(3)}L^2m_2[I_{\omega}](\Delta\rr)\, d\omega.\end{aligned}\]
If we set $m_2(\Delta \rr) = 0$ for $||\Delta\rr||>2L$, the above equality holds for all $\Delta\rr$, and taking its Fourier transform (with respect to $\Delta\rr$) and interchanging the Fourier integral with the integral over SO(3) (which can \emph{always} be done), we get
\[ \widehat{m_2[\II]}(\kk) \to \gamma\int_{\text{SO}(3)}L^2\widehat{m_2[I_{\omega}]}(\kk)\, d\omega = \gamma\langle|\widehat I_{\omega}(\kk)|^2\rangle_{\omega},\]
where $\langle\cdot\rangle_{\omega}$ denotes average over the orientations, so over SO(3).

@@

We derive a relation between the steered bispectrum derived in Sect.~\ref{sec:steering} to the volume, expanded in a suitable basis. Specifically, expand the Fourier-transformed volume as
\[ \widehat V(c\kk) = \sum_{\ell,m,s}a_{\ell,m,s}j_{\ell,s}(k)Y_{\ell,m}(\kk/k),\]
for $k\leq 1$ and zero otherwise, where $c$ is the assumed bandlimit, $Y_{\ell,m}$ may be taken to be either real or complex, and $j_{\ell,s}(r)$ is some radial basis function (in practice, either spherical bessel or the radial part of the 3D PSWFs as in [Lederman]). We work here with the complex spherical harmonics, given by
\[ Y_{\ell,m}(\theta,\varphi) = \sqrt{\frac{2\ell+1}{4\pi}\cdot\frac{(\ell-m)!}{(\ell+m)!}}P_{\ell}^m(\cos\theta)e^{i m\varphi},\]
where $P_{\ell}^m$ are the associated Legendre polynomials with the Condon-Shortley phase, and $j_{\ell,s}$ are normalized spherical Bessel functions given by
\[ j_{\ell, s}(k) = \frac{4}{|j_{\ell+1}(u_{\ell, s})|}j_{\ell}(2u_{\ell,s} k),\]
where $j_{\ell}$ is the spherical Bessel function of order $\ell$, $u_{\ell,s}$ is the $s$th positive zero of $j_{\ell}$, and the radius satisfies $k\in[0,1/2]$ in Fourier space since $1/2$ is the Nyquist frequency [ISBI].

@@

\begin{align*}
&C_2^{(q)} = \int_0^1j_{\ell,s_1}(k)j_{\ell,s_2}(k)R_{0,q}(k)k\, dk\\
&C_3^{(k,q_1,q_2)}\ell_1,m_1,s_1;\ell_2,m_2,s_2;\ell_3,s_3):= (-1)^{m_1+m_2}\left(\begin{array}{ccc}\ell_1 & \ell_2  & \ell_3\\ m_1 & m_2 & -m_1-m_2\end{array}\right)\\
&\times \sum_{N_1=-\ell_1}^{\ell_1}\sum_{N_2=-\ell_2}^{\ell_2}(-1)^{N_1+N_2}\left(\begin{array}{ccc}\ell_1 & \ell_2  & \ell_3\\ N_1 & N_2 & -N_1-N2\end{array}\right)\int_{\rr}\Psi_{\ell_1,N_1,s_1}(\rr)\rho_{\ell_2,N_2,s_2}^{(k,q_1)}(\rr)\overline{\rho_{\ell_3,N_1+N_2,s_3}^{(k,q_2)}(\rr)},
\end{align*}
and 
\begin{align*}
\rho_{\ell,N,s}^{(k,q)}&=\int_{\Delta\rr}\Psi_{\ell,N,s}(\rr+\Delta\rr)\overline{\psi_{k,q}(\Delta\rr)}.
\end{align*}
